---
title: "unintegrated_visualisation"
author: "Siwanart Ma"
date: "2025-03-20"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(Seurat)
library(tidyverse)
library(patchwork)
```

# Load the 3 Seurat objects
```{r}
# Increase the memory
mem.maxVSize(vsize = 60000)

# Load 4-6 weeks Seurat object
load("/Users/mayongzhi/Desktop/researchProject/integration/originals/human4_6W.RData")
ls()

# Load 7-11 weeks Seurat object
load("/Users/mayongzhi/Desktop/researchProject/integration/originals/human7_11W.RData")
ls()
human7_11W <- seurat
rm(seurat)

# Load 12-20 weeks Seurat object
load("/Users/mayongzhi/Desktop/researchProject/integration/originals/human12_20W.RData")
ls()
human12_20W <- humanpancreas.combined.sct
rm(humanpancreas.combined.sct)
```

# Visualisation
## human4_6W

The Seurat object is filtered by
nCount_RNA > 4000 & nCount_RNA < 50000 &
nFeature_RNA > 500 & nFeature_RNA < 8000 &
percent.mt < 15

Hence, we don't do filtering again.

```{r}
p <- VlnPlot(human4_6W, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), 
             ncol = 3, group.by = "days", pt.size = 0)

# Apply boxplot to each plot separately
p <- lapply(p, function(x) x + geom_boxplot(width = 0.1, outlier.shape = NA))

# Print the updated plots
patchwork::wrap_plots(p, ncol = 3) + theme_minimal()

```


Add a column "time" to mark object source for integration

```{r}
human4_6W$time <- "4_6"
```

## human7_11W

The Seurat object is filtered by
nCount_RNA > 4000 & nCount_RNA < 50000 &
nFeature_RNA > 500 & nFeature_RNA < 8000 &
percent.mt < 15

Hence, we don't do filtering again.

```{r}
# Reorder the 'days' variable to follow the desired order
human7_11W$days <- factor(human7_11W$days, levels = c("W7", "W8", "W9", "W10", "W11"))

p <- VlnPlot(human7_11W, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), 
             ncol = 3, group.by = "days", pt.size = 0)

# Apply boxplot to each plot separately
p <- lapply(p, function(x) x + geom_boxplot(width = 0.1, outlier.shape = NA))

# Print the updated plots
patchwork::wrap_plots(p, ncol = 3) + theme_minimal()

# Add a column "time" to mark object source for integration
human7_11W$time <- "7_11"
```
## human8_20W

The Seurat object is filtered by
nCount_RNA > 500 &
nFeature_RNA > 250 &
log10GenesPerUMI > 0.75 &
percent.mt < 25

```{r}
human12_20W<- subset(human12_20W, subset = nCount_RNA > 500 & nFeature_RNA > 250 & log10GenesPerUMI > 0.75 & percent.mt < 25)
```

```{r}
p <- VlnPlot(human12_20W, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), 
             ncol = 3, group.by = "Gestation_Age", pt.size = 0)

# Apply boxplot to each plot separately
p <- lapply(p, function(x) x + geom_boxplot(width = 0.1, outlier.shape = NA))

# Print the updated plots
patchwork::wrap_plots(p, ncol = 3) + theme_minimal()

# Add a column "time" to mark object source for integration
human12_20W$time <- "12_20"
```
## The dimension of each Seurat object
```{r}
print("human4_6W")
print(paste("Number of genes:", dim(human4_6W)[1], "; Number of cells:", dim(human4_6W)[2]))

print("human7_11W")
print(paste("Number of genes:", dim(human7_11W)[1], "; Number of cells:", dim(human7_11W)[2]))

print("human12_20W")
print(paste("Number of genes:", dim(human12_20W)[1], "; Number of cells:", dim(human12_20W)[2]))
```


## Unify the format of 3 Seurat objects

```{r}
table(human12_20W@meta.data$Gestation_Age)
```

According to Olaniru et al. 2023, there are 7168 cells in 13 PCW. Hence, we combine the cells from 13 PCW and 13 PCW_1

```{r}
human12_20W@meta.data$Gestation_Age[human12_20W@meta.data$Gestation_Age == "13PCW_1"] <- "13PCW"
```

```{r}
library(dplyr)

# Modify the "Gestation_Age" column in metadata
human12_20W@meta.data$Gestation_Age <- human12_20W@meta.data$Gestation_Age %>%
  as.character() %>%  # Ensure it's a character vector
  gsub("PCW", "", .) %>%  # Remove "PCW"
  paste0("W", .)  # Add "W" at the beginning

# Convert to a factor with a specific order
human12_20W@meta.data$Gestation_Age <- factor(human12_20W@meta.data$Gestation_Age, 
                                         levels = c("W12", "W13", "W14", 
                                                    "W15", "W18", "W19", "W20"))

# Set the new identity
Idents(human12_20W) <- human12_20W@meta.data$Gestation_Age

# Rename "Gestation_Age" with "days"
colnames(human12_20W@meta.data)[colnames(human12_20W@meta.data) == "Gestation_Age"] <- "days"
```


```{r}
four_six <- RunPCA(human4_6W, verbose = FALSE)
four_six <- RunUMAP(human4_6W, dims = 1:50, verbose = FALSE)

four_six <- FindNeighbors(human4_6W, dims = 1:50, verbose = FALSE)
four_six <- FindClusters(human4_6W, verbose = FALSE)
ggsave("../outputs/unintegrated_figure/UMAP4_6W.png",
       plot = DimPlot(four_six, label = TRUE) + ggtitle("UMAP for human4_6W"),
       width = 8, height = 6, dpi = 300)
DimPlot(four_six, label = TRUE) + ggtitle("UMAP for human4_6W")
```

```{r}
seven_eleven <- RunPCA(human7_11W, verbose = FALSE)
seven_eleven <- RunUMAP(human7_11W, dims = 1:50, verbose = FALSE)
seven_eleven <- FindNeighbors(human7_11W, dims = 1:50, verbose = FALSE)
seven_eleven <- FindClusters(human7_11W, verbose = FALSE)
ggsave("../outputs/unintegrated_figure/UMAP7_11W.png",
       plot = DimPlot(seven_eleven, label = TRUE) + ggtitle("UMAP for human7_11W"),
       width = 8, height = 6, dpi = 300)
DimPlot(seven_eleven, label = TRUE)+ ggtitle("UMAP for human7_11W")
```

```{r}
twelve_twenty <- RunPCA(human12_20W, verbose = FALSE)
twelve_twenty <- RunUMAP(human12_20W, dims = 1:20, verbose = FALSE)
twelve_twenty <- FindNeighbors(human12_20W, dims = 1:20, verbose = FALSE)
twelve_twenty <- FindClusters(human12_20W, verbose = FALSE)
ggsave("../outputs/unintegrated_figure/UMAP12_20W.png",
       plot = DimPlot(twelve_twenty, label = TRUE) + ggtitle("UMAP for human12_20W"),
       width = 8, height = 6, dpi = 300)
DimPlot(twelve_twenty, label = TRUE)+ ggtitle("UMAP for human12_20W")
```

# Import the merged seurat object for downstream visualisations
```{r}
library(ggplot2)

# Merge the Seurat objects
merged_object <- merge(x = human4_6W, y=list(human7_11W, human12_20W))
#saveRDS(merged_object, file = "../outputs/merged_object_13.rds")
merged_object <- readRDS("../outputs/merged_object_13.rds")
# Convert to a factor with a specific order
merged_object@meta.data$days <- factor(merged_object@meta.data$days, 
                                         levels = c("W4", "W5", "W6", 
                                                    "W7", "W8", "W9", "W10", "W11",
                                                    "W12", "W13", "W14", "W15", "W18", "W19", "W20"))

```


```{r}
# Plot violin plots
p <- VlnPlot(merged_object, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), 
             ncol = 1, group.by = "days", pt.size = 0) 

# Convert to list and apply boxplot separately
plots <- lapply(p, function(x) x + geom_boxplot(width = 0.1, outlier.shape = NA, show.legend = FALSE))

# Save the plots
for (i in seq_along(plots)) {
  ggsave(filename = paste0("../outputs/unintegrated_figure/plot_", i, ".png"), plot = plots[[i]], width = 8, height = 6, dpi = 300)
}

plots
```
## Bar plot

Visualize the number of cell counts in each period

```{r}
# Reorder the time in merged_object
merged_object$time <- factor(merged_object$time, levels = c("4_6", "7_11", "12_20"))

# Plot
plot <- merged_object@meta.data %>% 
  ggplot(aes(x=time, fill=time)) + 
  geom_bar(color="black") +
  stat_count(geom = "text", colour = "black", size = 3.5, 
            aes(label = ..count..),
            position=position_stack(vjust=0.5))+
  theme_classic() +
  theme(plot.title = element_text(hjust=0.5, face="bold")) +
  ggtitle("Number of Cells per Period")

# Save the plot
ggsave("../outputs/unintegrated_figure/nCellsPeriod.png", plot = plot,
       width = 8, height = 6, dpi = 300)

plot
```

```{r}
# Plot
plot <- merged_object@meta.data %>% 
  ggplot(aes(x=days, fill=days)) + 
  geom_bar(color="black") +
  stat_count(geom = "text", colour = "black", size = 3.5, 
            aes(label = ..count..),
            position=position_stack(vjust=0.5), angle = 90)+
  theme_classic() +
  theme(plot.title = element_text(hjust=0.5, face="bold")) +
  ggtitle("Number of Cells per Week")

# Save the plot
ggsave("../outputs/unintegrated_figure/nCellsWeek.png", plot = plot,
       width = 8, height = 6, dpi = 300)

plot
```
## Density Plot

Visualise the distribution of nCount, nFeature, and percent.mt between each seurat object
```{r}
p <- merged_object@meta.data %>% 
    ggplot(aes(color=time, x=nCount_RNA, fill= time)) + 
    geom_density(alpha = 0.2) + 
    theme_classic() +
    scale_x_log10()

ggsave("../outputs/unintegrated_figure/den_C.png", plot = p, width = 8, height = 6, dpi = 300 )
p
```

```{r}
p <- merged_object@meta.data %>% 
    ggplot(aes(color=time, x=nFeature_RNA, fill= time)) + 
    geom_density(alpha = 0.2) + 
    theme_classic() +
    scale_x_log10()
ggsave("../outputs/unintegrated_figure/den_F.png", plot = p, width = 8, height = 6, dpi = 300 )
p
```

```{r}
p <- merged_object@meta.data %>% 
    ggplot(aes(color=time, x=percent.mt, fill= time)) + 
    geom_density(alpha = 0.2) + 
    theme_classic() +
    scale_x_log10()

# geom_density: creaty a density plot. alpha = transparency
# scale_x_log10: Transforms the x-axis to a logarithmic scale (base 10), so the large values are compressed. Good for visualizing skewed data

ggsave("../outputs/unintegrated_figure/den_mt.png", plot = p, width = 8, height = 6, dpi = 300 )
p
```

## Feature Scatter Plot

Scatter plot of percent.mt and nFeature

```{r}
p <- FeatureScatter(merged_object, feature1 = "percent.mt", feature2 ="nFeature_RNA" , group.by="orig.ident",split.by="time", ncol = 3) 

ggsave("../outputs/unintegrated_figure/FeaSca_mtF.png",
       plot = p,
       width = 12, height = 6, dpi = 300) 
p
```

```{r}
p <- FeatureScatter(merged_object, feature1 = "percent.mt", feature2 ="nFeature_RNA" , group.by="days",split.by="time", ncol = 3) 

ggsave("../outputs/unintegrated_figure/FeaSca_mtF_days.png",
       plot = p,
       width = 12, height = 6, dpi = 300) 
p
```
Scatter plot of nCount and nFeature

```{r}
p <- FeatureScatter(merged_object, feature1 = "nCount_RNA", feature2 ="nFeature_RNA" , group.by="orig.ident",split.by="time", ncol = 3) 

p <- p + theme(axis.text.x = element_text(angle = 45, hjust = 1))
          
ggsave("../outputs/unintegrated_figure/FeaSca_CF.png",
       plot = p,
       width = 12, height = 6, dpi = 300) 
p
```

```{r}
p <- FeatureScatter(merged_object, feature1 = "nCount_RNA", feature2 ="nFeature_RNA" , group.by="days",split.by="time", ncol = 3) 

p <- p + theme(axis.text.x = element_text(angle = 45, hjust = 1))
          
ggsave("../outputs/unintegrated_figure/FeaSca_CF_day.png",
       plot = p,
       width = 12, height = 6, dpi = 300) 
p
```


